Why robot programming is hard

### 2013-07-18 paper.tex
Done for the day. Still need to revisit the Concurrency section, but I've made a full pass over the rest.

A different aspect of high-dimensionality comes from the design of robot arms. A solid object in 3D space has six degrees of freedom (DOF): three dimensions to specify its position, and three more to specify its rotation. If a robot arm had only 6-DOF and was trying to execute a specific grasp of a particular object, such as picking up soda can from the side, there would be only a single arm configuration that would allow the gripper to pick up the can. In contrast, human arms and advanced robot arms have 7-DOF, which allows them one extra dimension of flexibility in achieving those grasps. To visualize this, imagine grasping a soda can on the table in front of you and then flapping your elbow up and down while keeping the can stationary. That extra degree of freedom enables your arm to take on all those different shapes while still achieving the desired grasp.

Interfaces for robot arm teleoperation typically involve setting the position or angle of each joint in the arm individually. For example, a tablet-style interface might have fourteen buttons, one pair used to increase or decrease the angle of each joint in the arm. That additional degree of freedom makes it easier to identify a valid arm configuration for executing a grasp, since only one of many valid configurations must be found, not the unique one.

For teleoperators who are manually positioning a robot arm must only find one way to position the various joints

It may also make it easier for teleoperators to find a valid arm configuration to reach a desired grasp. Yet those seven degrees of freedom also represent seven variables to control when using a direct-manipulation teleop interface

High dimensionality also makes it difficult to specify to how pick up objects using a robotic gripper. A solid object has six degrees of freedom: three degrees to specify its position, and another three to specify its rotation. The human arm, and the most advanced robotic arms, have seven degrees of freedom.  Many robot arm control interfaces ask users to independently specify joint angles for each of those seven degrees of freedom. These interfaces are cumbersome because the joints are interdependent: setting the angle of one joint, such as the elbow, dictates what angles are possible for the shoulder and vice versa. There is a tradeoff between exposing these joint limitations to the programmer, which allows precise yet complex control, versus insulating a programmer from those considerations, which may lead to unpredictable behavior that is difficult to debug.

Expert roboticists manage high-dimensionality by representing the world in terms of {\em reference frames} (coordinate systems that are defined relative to particular robot part, such as the base or the head-mounted camera) and use matrix algebra to convert between different frames of reference. Robot joint positions are typically specified relative to their own reference frame; for example, a robot's wrist, which is located at the end of its arm, can be commanded to rotate around the axis that is aligned with the robot's forearm (similar to the motion we use to screw in a light bulb). Calculating the absolute position of a robot's hand (for example, to determine whether it is near the door handle) requires a series of complex matrix calculations starting from the position of its base, through its shoulder, and all the joints, down to the hand. The opposite problem, calculating the joint positions required to get the hand to a particular position, is called {\em inverse kinematics} and is typically solved using either analytical equations or heuristic approximation methods.

Toolkits for doing automatic motion planning are on the verge of becoming mature, simplifying the problem of robot arm control from how the arm should move in fine-grained detail to specifying only the high-level target position and orientation. Yet the challenge of specifying that 6-DOF target position and orientation still remains, and this will continue to be a challenge for easy-to-use robot interfaces that work with existing 2D displays and input devices.

@@

What does this mean for robot programmers? Robot programs must have mechanisms in place to deal with uncertain truth and unpredictable hardware. Robot control loops usually sense the physical hardware during control to determine whether a motion was carried out successfully. 
% TODO
Rather than simplying instructing the robot to pick up an object, the program must specify what to do if the object is not {\em where} it thinks it is, is not {\em what} it thinks it is, or if the robot fails to perform the desired grasp.

In order to deal with these various types of uncertainty, professional programmers can learn to use probabilistic models such as Bayesian networks or HMMs, or they write code with lots of built-in error checking. However, casual programmers who just want a robot to perform a job in the home would find it burdensome to have to elaborate recovery strategies for each of the different failure conditions.

Most end user programming systems to date ignore the problem of uncertainty. Some can get away with this because they operate in a simulated world (such as Scratch). Others, which deal with uncertainty in the form of network requests timing or webservices being down, rely on the user to retry the operation as needed. A good user interface paradigm for handling uncertainty will be needed for end user programmed robot control software to be robust.

@@

This past generation of commercial home robots illustrates one extreme end of the simplicity/expressiveness tradeoff. iRobot's successful Roomba line of vacuum-cleaning robots\footnote{http://www.irobot.com/us/learn/home/roomba.aspx} sports a single-button interface that simply says ``Clean''.

The next generation of commercial robots in this space will likely look like Rethink Robots' Baxter system\footnote{http://www.rethinkrobotics.com/index.php/products/baxter/}.  Baxter is a two-armed manipulator than can be programmed by demonstration to automate factory work simply by guiding its arms to the appropriate position and tapping icons on the robot's ``face''.  This style of programming makes it possible for people to specify the robot's behavior directly on the robot, without having to deal with any abstractions. In addition, Baxter performs its work in a human-safe way, immediately ceasing motion if it detects that a human has moved into its path, enabling it to work side by side with factory workers unlike previous generations of industrial robots.

For the future, our team at Willow Garage is exploring different tradeoff points in the space of simplicity versus expressiveness. One research direction we have been exploring is to look at specific verticals and design a customized set of robot programming primitives specifically for that domain. Each of these primitives can be hand-crafted in the lab by expert roboticists to embody a rich amount of automation, parallelism, and support for uncertainty. The end user programming problem is then a simpler matter of sequencing these higher-level primitives into useful, customizable programs for their environment.

For example, our team has been doing ethnographic investigations of employees in continuing care retirement centers to discover how robotic helpers could enable them to spend more time with residents. We have discovered that a robotic assistant who could autonomously fetch and deliver supplies to arbitrary locations within the facility could be beneficial to their workflow. Based on those findings, we have defined a set of primitives that include picking up and dropping off supply bins, navigating autonomously to pre-defined locations, and prompting for user input via a touchscreen on the robot's ``face''. End user programmers will be able to sequence these primitives themselves to design custom behaviors for a robotic delivery system, while expert roboticists can focus on making individual primitives robust using the full slate of advanced programming techniques at their disposal.

Over time, as we discover and build new domain-specific primitives, my hope is that we will expand the repertoire of robust end-user-programmable robot behavior. And perhaps that goal of having our own Rosie the Robot in our homes will soon become a reality.

>>>>

A different aspect of high-dimensionality comes from the design of robot arms. A solid object in 3D space has six degrees of freedom (DOF): three dimensions to specify its position, and three more to specify its rotation. If a robot arm had only 6-DOF and was trying to execute a specific grasp of a particular object, such as picking up soda can from the side, there would be only a single arm configuration that would allow the gripper (a robotic hand) to pick up the can. In contrast, human arms and advanced robot arms have 7-DOF, which allows them one extra dimension of flexibility in achieving those grasps. To visualize this, imagine grasping a soda can on the table in front of you and then flapping your elbow up and down while keeping the can stationary. That extra degree of freedom enables your arm to take on all those different shapes while still achieving the desired grasp.

Interfaces for robot arm teleoperation typically involve setting the position or angle of each joint in the arm individually. For example, a tablet-style interface might have fourteen buttons, one pair used to increase or decrease the angle of each joint in the arm. That additional degree of freedom makes it easier to identify a valid arm configuration for executing a grasp, since only one of many valid configurations must be found, not the unique one.  However controlling each individual joint position by clicking on buttons is tedious. Additionally, since the allowable joint positions are not completely independent, finding a correct solution is often done through trial and error.

%There is a tradeoff between exposing these joint limitations to the programmer, which allows precise yet complex control, versus insulating a programmer from those considerations, which may lead to unpredictable behavior that is difficult to debug.

Yet positioning an arm through direct manipulation or teleop is relatively easy compared to doing it programmatically. In order to program an arm, one must first specify {\em where} it should go.  Expert roboticists represent the world in terms of {\em reference frames} (coordinate systems that are defined relative to a fixed position). For example, one reference frame could be relative to the building, with the front door being at coordinate (0, 0).  Other reference frames are defined relative to stationary parts of the robot, such as its base (if it is on wheels), the head-mounted camera, or the right elbow joint.

Sending an arm to a specific position is useful for robot tasks such as, for example, looking at what it is holding by moving its hand to half a meter in front of its face. This half-meter is most reliably specified in the reference frame relative to the robot's head. Given this goal position, the programmer must find a series of arm configurations that moves the arm to the desired place, while avoiding contact with obstacles in the environment.

% such as the base or the head-mounted camera) and use matrix algebra to convert between different frames of reference. Robot joint positions are typically specified relative to their own reference frame; for example, a robot's wrist, which is located at the end of its arm, can be commanded to rotate around the axis that is aligned with the robot's forearm (similar to the motion we use to screw in a light bulb). Calculating the absolute position of a robot's hand (for example, to determine whether it is near the door handle) requires a series of complex matrix calculations starting from the position of its base, through its shoulder, and all the joints, down to the hand. The opposite problem, calculating the joint positions required to get the hand to a particular position, is called {\em inverse kinematics} and is typically solved using either analytical equations or heuristic approximation methods.

Toolkits for doing automatic motion planning are maturing, which will simplify the problem of robot arm control from how the arm should move in fine-grained detail to specifying only the high-level target position and orientation. Yet the challenge of specifying that 6-DOF target position and orientation still remains, and this will continue to be a challenge for easy-to-use robot interfaces that work with existing 2D displays and input devices.

@@

What does this mean for robot programmers? Robot programs must have mechanisms in place to deal with uncertain truth and unpredictable hardware. Robot control loops usually sense the physical hardware during operation to determine whether a motion was carried out successfully. Unlike with demonstration and teleop, where the operator can see if the robot carried out the desired motion and take appropriate action, a robot programmer must think through possible outcomes while writing a program and craft appropriate responses for each eventuality. For example, while instructing a robot to pick up an object, a programmer must specify what to do if the object is not {\em where} it thinks it is, is not {\em what} it thinks it is, or if the robot fails to perform the desired grasp. Should it simply try again? Move a little bit to get a better view and then try again? Ask a human?

In order to deal with these various types of uncertainty, professional programmers learn to use probabilistic models and learn to craft fault-tolerant code with frequent error checking, which seems like an heavy burden for casual programmers. Most end user programming systems to date simply ignore the problem of uncertainty. For robot programming to be successful, we will need to develop easily-understandable paradigms for handling uncertainty.

@@

This past generation of commercial home robots illustrates one extreme end of the simplicity versus expressiveness tradeoff. iRobot's successful Roomba line of vacuum-cleaning robots\footnote{http://www.irobot.com/us/learn/home/roomba.aspx} sports a single-button interface that simply says ``Clean''. While the robot will eventually clean your floors, you lose control over how exactly it does so, as it performs what seems like a random walk through your home.

%The next generation of commercial robots in this space will likely look like Rethink Robots' Baxter system\footnote{http://www.rethinkrobotics.com/index.php/products/baxter/}.  Baxter is a two-armed manipulator than can be programmed by demonstration to automate factory work simply by guiding its arms to the appropriate position and tapping icons on the robot's ``face''.  Unlike past demonstration-based robots which mindlessly repeated the same motions, Baxter has an intelligent motion planner which can immediately cease moving the arms if it detects something in the way, such as a human walking into its workspace.

%This style of programming makes it possible for people to specify the robot's behavior directly on the robot, without having to deal with any abstractions. In addition, Baxter performs its work in a human-safe way, immediately ceasing motion if it detects that a human has moved into its path, enabling it to work side by side with factory workers unlike previous generations of industrial robots.

Roboticists are hard at work developing increasingly autonomous robot capabilities for well-specified tasks such as 2D navigation (getting from point A to point B within a building), arm motion planning (moving the gripper to a specified position and orientation while avoiding obstacles), and recognizing and picking up discrete objects on a flat surface. Autonomy is key for enabling robots to automatically perform longer and more complex tasks on their own, but it also reduces expressivity by limiting the ways in which those tasks can be performed to those that are programmed in to the automation.

Our approach to designing end-user-programmable interfaces is to craft these autonomous components into a good set of {\em primitives} in an end user programming language, such that a non-expert would not have to ``look inside the box'' to understand how each one works. Roboticists can then implement these primitives using advanced techniques to solve the difficult problems encountered during robot programming. For example, a good autonomous navigation component could incorporate a set of recovery behaviors to try on navigation failure, such as turning around, waiting for people/obstacles to clear out of its path, and backing up and retrying. It could also bake in some concurrent strategies, such as turning the head to look in the direction the robot is going.

%Encapsulating all this functionality into a single component with a well-defined interface makes it easier for end user robot programmers by giving them more reliable, easier to understand components they can assemble into programs.

We have made use of these autonomous functions to develop a system at Willow Garage called Continuous Ops that enables casual users to control our PR2 robots through a web interface (Figure~\ref{contops}). This web interface, accessible from standard laptops or tablets, provides the ability to tell a PR2 robot to unplug itself from the power supply, navigate autonomously to anywhere in the buliding, position itself in front of a table, automatically identify graspable objects on that table, and pick up an object.

This interface lies somewhere between teleoperation and programming. It does not require an operator to exert direct control over the robot, but the commands that are given are executed immediately by the robot rather than scheduled to occur later in time. We are also exploring interfaces for sequencing these primitives into scheduled programs, in order to validate that the set of primitives we are developing is complete enough to perform useful tasks.

Another key component of our approach, which we call {\em shared autonomy}, is to fall back on teleoperator assistance when autonomy fails, as it will invariably do for the near future. In our system designs, the teleoperator does not necessarily have to be the person who wrote the program. Instead, we are modeling our system after telephone call centers where a team of trained professionals handle incidents as they arise and get the robots back on track to fulfill their programmed goals. As a stepping-stone towards full autonomy, this approach enables us to start testing end user programmable systems without needing a fully reliable set of autonomous primitives in place to support them.

Over time, as we refine existing autonomous primitives and build new domain-specific primitives, my hope is that we will expand the repertoire of robust end-user-programmable robot behavior. And perhaps that goal of having our own Rosie the Robot in our homes will soon become a reality.
